---
title: Unit tests for KOM functions
date: 2017-08-26
author: Bret Beheim
---

There's a lot of complicating operations in KOM, so it is important to unit test every function, constantly, trying out scenarios well beyond the ones you initially developed them for.

The two important functions are `neighbor_finder` and `kin_retriever`, which are used to detect alters who are relationally connected by spatial location or by kinship.

We can illustrate how these work by first simulating a population.

```{r, eval=FALSE}

pop <- simanize()

```

I'll sample one individual at `target_pid`. We pass this ID 

```{r, eval=FALSE}

neighbor_finder( target_pid, data = pop)

```

The output is all the people who live nearby. We can expand this search area by the `dist_radius` argument, which is in meters.

Likewise with `kin_retriever`, 

```{r, eval=FALSE}

kin_retriever( target_pid, data = pop)

```

This information is useful because, for example, now we can look up those people and see if they have the diffusing trait.

In the ./tests folder, you will find a series of unit tests for these functions, that pass a variety of different inputs and check that the results are right.

Of course these tests are just for me! They aren't included in the downloadable version of the package.

Finally there's the function `pequi_get` which is how diffusing traits are generated. This of course needs to be renamed to trait_get or soemthing...

